---
  output: github_document
---
  
  <!-- README.md is generated from README.Rmd. Please edit that file -->
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sandwich2stage

<!-- badges: start -->
<!-- badges: end -->

`sandwich2stage` introduces a function for computing an estimate of the sandwich variance for the two-stage regression model setting of regression calibration. The sandwich is one approach for obtaining standard errors in two-stage regression settings that account for the extra uncertainty added by the calibration model step. Our function computes an estimate of the sandwich variance obtained by stacking the stage 1 and stage 2 estimating equation contributions.  

## Installation

`sandwich2stage` may be installed from github as follows:  

``` {r}
library(devtools)
install_github("lboe23/sandwich2stage", subdir="pkg")
```

## Example

After we have loaded our package, we will want to load in the sample data assumed to be from a simple random sample, which is contained in our package.

```{r load data}
library(sandwich2stage)
data("sandwichdata_SRS")
```
Next, we need to load the survey package and create a survey design object. For the case of the data `sandwichdata_SRS` from a simple random sample, we will want to specify a simple random sampling design.

```{r create design}
library(survey)
sampdesign <- svydesign(id=~1, data=sandwichdata_SRS)
```
We may then fit the stage 1 and stage 2 models, saving the estimated nuisance parameters from the stage 1 model and using them to obtain an estimated of the unknown exposure (xhat). This estimate exposure, xhat, will be used as a covariate in the stage 2 model. Note that the stage 1 model is only fit to the subset which contains validation data (i.e. where v=1).

```{r stage 1}
stage1.model<-survey::svyglm(xstarstar~xstar+z,design=sampdesign,family=gaussian(),subset=v==1)
alphas.stage1<-coef(stage1.model)
sampdesign <- update(sampdesign,xhat =predict(stage1.model,newdata=sampdesign$variables) )
```

We will then fit the stage 2 model with the estimated exposure (xhat) as a covariate.

```{r stage 2}
stage2.model<-  survey::svyglm(y ~ xhat+z,design=sampdesign,family=binomial())
```

Finally, we can obtain an estimate of the sandwich variance using our function`sandwich2stage()`. The sandwich variance matrix is saved below in the object `sandwichvar`. 

```{r sandwich} 
sandwich.object<-sandwich2stage(stage1.model,stage2.model,xstar="xstar",xhat="xhat",Stage1ID="ID",Stage2ID="ID")
sandwichvar<-vcov(sandwich.object) 
```

