---
title: "Sandwich Variance in 2 Stage Regression with Data from a Complex Survey Design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sandwich Variance in 2 Stage Regression with Data from a Complex Survey Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of the package sandwich2stage is to compute an estimate of the sandwich variance for two-stage regression settings. Our function assumes that at stage 1, a linear regression calibration model is fit to obtain a plug-in estimate of the nuisance parameters, then at stage 2, the user fits an outcome model that relies on these plug-in estimates. The user must first load our package, as follows: 

```{r setup}
library(sandwich2stage)
```

First, we load the data from our package. Note that this sample data set is assumed to be from a complex survey design with stratification, clustering, and unequal probability weighting. This data set has XXX rows (e.g. subjects) and 9 columns:
- $ID$: a unique ID variable for each subject in the data frame
- $PSUid$: the primary sampling unit (PSU) or cluster ID variable
- $strat$: the variable specifying the strata of the survey design
- $myweights$: the sampling weights reflecting unequal probability of selection into the sample
- $xstar$: the error-prone, continuous exposure variable, available on all cohort study subjects (e.g. self-reported energy intake)
- $xstarstar$: the exposure variable prone to classical measurement error, available only on a subset (e.g. recovery biomarker for energy intake)
- $v$: an indicator variable for whether subjecs are in the subset used to fit the stage 1 model
- $z$: a precisely-measured continuous covariate in the data set, also assumed to be related to the outcomes of interest
- $y$: a binary outcome variable in the data set (e.g. hypertension) to be used in a logistic regression model
- $Time$: failure time in years, to be used in a Cox model
- $delta$: an indicator of whether failure or right-censoring occured at the end of the time period, to be used in a Cox model

```{load data}
data("sandwichdata_svy")
head(sandwichdata_svy)
```

Next, we will create the survey design object, specifying that this is a complex survey design. We begin

```{svy design}
sampdesign <- svydesign(id=~PSUid, strata=~strat,weights=~myweights, data=sandwichdata_svy)  
```
Now we fit the stage 1 and stage 2 models, saving the estimated nuisance parameters from the stage 1 model and using them to obtain an estimated of the unknown exposure ($\hat{X}_i$), which will be used in the stage 2 model. We update the survey design to include this new estimated exposure (``xhat"). We will first assume that we are interested in fitting a logistic regression stage 2 outcome model using the binary outcome variable $y$ using svyglm() with a binomial family.

```{fit models}
stage1.model<-svyglm(xstarstar~xstar+z,design=sampdesign,family=gaussian(),subset=v==1)
alphas.stage1<-coef(stage1.model)
sampdesign <- update(sampdesign,xhat =predict(stage1.model,newdata=sampdesign$variables) )
stage2.model.glm<-  svyglm(y ~ xhat+z,design=sampdesign,family=binomial())
```

Finally, our function sandwich2stage() may be used to obtain an estimate of the sandwich variance. This function requires the following arguments: the fitted stage 1 and 2 model objects, the names of the variables $X^*_i$ and $\hat{X}_i$ in the data, and the names of the subject ID variables for stage 1 and stage 2 models. 


```{obtain sandwich 1}
sandwich.object1<-sandwich2stage(stage1.model,stage2.model.glm,
                                  xstar="xstar",xhat="xhat",
                                  Stage1ID="ID",Stage2ID="ID")
sandwichvar1<-vcov(sandwich.object1)
```

Now, suppose we are also interested in using the same estimated exposure variable $\hat{X}_i$ in a Cox proportional hazards stage 2 outcome model. Below, we fit this stage 2 model using svycoxph():

```{fit Cox model}
stage2.model.Cox<- svycoxph(Surv(Time, delta) ~ xhat+z, design=sampdesign)
```

Once again, we can use our function sandwich2stage() to obtain an estimate of the sandwich variance when the stage 2 model is a Cox model:

```{obtain sandwich 2}
sandwich.object2<-sandwich2stage(stage1.model,stage2.model.Cox,
                                  xstar="xstar",xhat="xhat",
                                  Stage1ID="ID",Stage2ID="ID")
sandwichvar2<-vcov(sandwich.object2)
```
